---
title: "CKME136"
author: "Chun Chow"
date: "2020/7/5"
output:
  word_document: default
  html_document: default
---

##Preparation and Data Import

#loading libraries 
#===============================================================
```{r libraries}

library(assertive)
library(dplyr)
library(stringr)
library(tidyverse)
library(ggplot2)
library(forcats)
library(VIM) 
library(forcats)
library(scales)
library(broom)
library(GGally)
library(tibble)
library(gridExtra)

```


#Data Preprocessing
#================================================================

#Import data
```{r new_car}
new_car <- read.csv("C:\\Users\\Justin\\Documents\\data\\new_car\\USA_cars_datasets.csv")

```

#Overview of data
#==============================================================
```{r}
str(new_car)
head(new_car)
summary(new_car)
```

#Check if there is any missing data
```{r}

sum(is.na(new_car))
table(is.na(new_car))
sum(complete.cases(new_car))
```

#Check is there any full duplicates in the data
```{r}
sum(duplicated(new_car)) 
```

#Checking levels on categorical data
```{r}
table(new_car$brand)
table(new_car$model)
table(new_car$title_status)
table(new_car$color) #too many different color, may need clear later
table(new_car$state)
table(new_car$country)
table(new_car$condition)
```

#Data Cleaning 
#===========================================================

```{r}
#Remove unneeded columns, the x, vin and condition columns does not provide useful information to the analysis
new_car[, c("X", "vin", "condition", "lot")] <- list(NULL)
names(new_car)
```

#Change year column into factor
```{r}
new_car <- new_car %>%
  mutate(year = as.factor(year))
str(new_car)
```

#Basic look at all the color levels of the car
```{r}
ggplot(new_car, aes(color)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 90))

levels(new_car$color)
```

#Too many different color name with slightly different, may want to group some of the similar color into one category
```{r}
yellow <- c("beige", "triple yellow tri-coat", "yellow")
red <- c("burgundy", "cayenne red", "maroon", "red", "royal crimson metallic tinted clearcoat", "ruby red", 
         "ruby red metallic tinted clearcoat", "toreador red")
grey <- c("charcoal", "gray", "magnetic metallic", "shadow black" )
black <- c("black", "black clearcoat", "phantom black", "super black", "tuxedo black metallic")
silver <- c("billet silver metallic clearcoat", "ingot silver" , "ingot silver metallic", "silver")
blue <- c("blue","dark blue", "jazz blue pearlcoat", "kona blue metallic", "light blue", "lightning blue", "morningsky blue", "turquoise")
white <- c("bright white clearcoat", "glacier white", "off-white", "oxford white", "pearl white", "white", 
           "white platinum tri-coat metallic")
brown <- c("brown", "tan" )

orange <- c("competition orange", "orange"  )
gold <- c("gold")
green <- c("green" )
purple <- c("purple")

#New col name "color_collapsed" for the collapsing categories
new_car2 <- new_car %>% 
  mutate(color_collapsed = fct_collapse(color, Yellow = yellow, Red = red, Grey = grey , Black = black, Silver = silver, Blue = blue, White = white, Brown = brown, Orange = orange, Gold = gold, Green = green , Purple = purple )) 

```

#From 49 different color categories into just 14 categories
```{r}
new_car %>% count(color)

new_car2 %>% count(color_collapsed)

table(new_car2$color_collapsed)
```

#Some of the color is undefined in the data
```{r}
#61 cars label as "no_color"
new_car2 %>%
  filter(color == "no_color")

#5 cars label as "color:"
new_car2 %>%
  filter(color == "color:")

#one vehicle label as "guard" color
new_car2 %>%
  filter(color == "guard")
```

#Put them into not_identify
```{r}
not_identify <- c("color:", "no_color", "guard")

new_car_collapsed <- new_car2 %>%
  mutate(color_collapsed = fct_collapse(color_collapsed, Not_identify = not_identify))
         
```

#Total factors in the color column
```{r}
levels(new_car_collapsed$color_collapsed)
table(new_car_collapsed$color_collapsed)

```

#Turn not_identify value into NA
```{r}
new_car_collapsed <- new_car_collapsed %>%
  mutate(color_collapsed = recode_factor(color_collapsed, "Not_identify" = NA_character_))


sum(is.na(new_car_collapsed$color_collapsed))
new_car_collapsed %>%
  filter(is.na(color_collapsed))

new_car_collapsed[, "color"] <- list(NULL)
```

#Use kNN imputation for the categorical NA value in "color_collapes" column
```{r}
new_car_collapsed_knn <- kNN(new_car_collapsed, variable = c("color_collapsed"), k =5)
new_car_collapsed_knn[,"color_collapsed_imp"] <- list(NULL)

sum(is.na(new_car_collapsed_knn$color_collapsed))

knn_before <- table(new_car_collapsed$color_collapsed)
knn_before <- as.data.frame(as.list(knn_before))

knn_after <- table(new_car_collapsed_knn$color_collapsed)
knn_after <- as.data.frame(as.list(knn_after))

knn_compare <- rbind(knn_before, knn_after)
rownames(knn_compare) <- c("Before kNN", "After kNN")
str(new_car_collapsed_knn)
knn_compare
```




#Identify the outliers on the Price and Mileage Column
```{r}
price_before <-ggplot(new_car_collapsed_knn, aes(x = 0,y = price)) +
  geom_boxplot()
mileage_before <- ggplot(new_car_collapsed_knn, aes(x = 0, y = mileage)) +
  geom_boxplot()
grid.arrange(price_before,mileage_before, nrow = 1)
```

#Remove outliers on the Price and Mileage Column
```{r}
Q_price <- quantile(new_car_collapsed_knn$price, probs = c(0.25, 0.75))
iqr_price <- IQR(new_car_collapsed_knn$price)

Q_mileage <- quantile(new_car_collapsed_knn$mileage, probs = c(0.25, 0.75))
iqr_mileage <- IQR(new_car_collapsed_knn$mileage)

new_car_outliers_removed <- new_car_collapsed_knn %>%
  filter(price >(Q_price[1] - 1.5 * iqr_price) & price < (Q_price[2] + 1.5 * iqr_price) &
         mileage >(Q_mileage[1] - 1.5 * iqr_mileage) & mileage < (Q_mileage[2] + 1.5 * iqr_mileage) )
```


#New dataset after removed outliers
```{r}
price_after <- ggplot(new_car_outliers_removed, aes(x = 0,y = price)) +
  geom_boxplot()

mileage_after <- ggplot(new_car_outliers_removed, aes(x = 0, y = mileage)) +
  geom_boxplot() 
  

grid.arrange(price_after, mileage_after, nrow = 1)
```

#Exploratory Data Analysis
#===========================================================================
#Visualize the data

#after cleaning graph
```{r}
ggplot(new_car_outliers_removed, aes(color_collapsed)) +
  geom_bar(stat = "count", aes(fill =  color_collapsed)) +
  scale_fill_manual(values = c("#FFCC33", "#CCCCCC", "#000000", "#0000FF", "cornsilk1", "#CC6600", "#990000", "#333333", "#FFCC00", "#FFCC66", "#66CC00", "#990099")) +
  ggtitle("Total Count of All Color") +
  theme_bw()
```

#Most of the cars are white in color
```{r}
new_car_outliers_removed %>%
  group_by(color_collapsed) %>%
  summarize(count = n()) %>%
  arrange(desc(count))
```


#Distribution of the price
```{r}
ggplot(new_car_outliers_removed, aes(price)) +
  geom_density(fill = "skyblue") +
  scale_x_continuous(breaks = seq(min(new_car_outliers_removed$price), max(new_car_outliers_removed$price), by = 5000)) + 
  ggtitle("Price Density")
```

#Information about car brand and their count
```{r}
ggplot(new_car_outliers_removed, aes(fct_infreq(brand), fill = brand)) +
  geom_histogram(stat = "count", bidwidth = 30) +
  labs(x = "Brand", y = "Total Count", title = "Total Count of All Brand") +
  theme(axis.text.x = element_text(angle = 0)) +
  coord_flip()

new_car_outliers_removed %>%
  group_by(brand) %>%
  summarize(count = n()) %>%
  arrange(desc(count))
  
```

#Information about the car model and their count
```{r}

new_car_outliers_removed %>%
  group_by(model) %>%
  summarize(count = n()) %>%
  arrange(desc(count))

  
new_car_outliers_removed %>%
  count(model) %>%
  ggplot(aes(x= reorder(model, n, sum), y= n, fill = model)) +
  geom_col() + 
  labs(x= "Model", y = "Total", title = "Total count") +
  theme(legend.position = "none") +
  coord_flip()
```

#The distribution of Mileage and Price
```{r}
ggplot(new_car_outliers_removed, aes(mileage, price)) +
  geom_point(alpha =.6) +
  geom_smooth(method = "lm", se = FALSE) +
  scale_x_continuous(breaks = seq(min(new_car_outliers_removed$mileage), max(new_car_outliers_removed$mileage), by = 50000)) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Mileage & Price")
```

#The distribution of Brand and Price
```{r}
dfcount <- count(new_car_outliers_removed, brand)

ggplot(new_car_outliers_removed, aes(brand, price)) +
    geom_boxplot(fill='#A4A4A4', color="black") +
  scale_y_continuous(breaks = seq(min(new_car_outliers_removed$price), max(new_car_outliers_removed$price), by = 2500)) +
  theme(axis.text.x = element_text(angle = 90)) +
    geom_text(data = dfcount, aes(y = 0, label = n))
```

#What is the most expensive car model for each brand
```{r}
most_expensive <- new_car_outliers_removed %>%
  group_by(brand) %>%
  top_n(1,price) %>%
  arrange(desc(price)) %>%
  select(brand, price, model)  
most_expensive
```

#The distribution of Year and Price
```{r}
give.n <- function(x){
  return(c(y = median(x)*1.08, label = length(x))) 
}
ggplot(new_car_outliers_removed, aes(year, price)) +
  geom_boxplot(fill='#A4A4A4', color="black") +
  theme(axis.text.x = element_text(angle = 90))
```

#The distribution of Color & Price
```{r}
ggplot(new_car_outliers_removed, aes(color_collapsed, price)) +
  geom_boxplot(fill='#A4A4A4', color="black") +
  stat_summary(fun.data = give.n, geom = "text", fun.y = median)  
```
#The distribution of State & Price
```{r}
ggplot(new_car_outliers_removed, aes(state, price)) +
  geom_boxplot(fill='#A4A4A4', color="black") +
  theme(axis.text.x = element_text(angle = 90)) +
  stat_summary(fun.data = give.n, geom = "text", fun.y = median)  
```

#The distribution of Vehicle statue & Price
```{r}
ggplot(new_car_outliers_removed, aes(title_status, price)) +
  geom_boxplot(fill='#A4A4A4', color="black") +
  stat_summary(fun.data = give.n, geom = "text", fun.y = median) 
```

#Correlation Test
#===============================================================
```{r}
#Check the distribution of the data
ggplot(new_car_outliers_removed, aes(price)) +
  geom_histogram(bins = 50)


#Pearson Correlation
new_car_outliers_removed %>%
  summarize(Total = n(), r = cor(mileage, price))

#Spearman Correlation
new_car_outliers_removed %>%
  summarize(Total = n(), r = cor(mileage, price, method = "spearman"))
#Becuase the data is not normal, spearman correlation was used for the correlation test
```


#Linear Regression
#======================================================================

#Filter out the very low counts levels for better regression analysis
```{r}
new_car_outliers_removed <- new_car_outliers_removed  %>%
  group_by(brand) %>%
  filter(n() > 5) %>%
  droplevels()

new_car_outliers_removed <- new_car_outliers_removed  %>%
  group_by(model) %>%
  filter(n() > 5) %>%
  droplevels()

new_car_outliers_removed <- new_car_outliers_removed  %>%
  group_by(color_collapsed) %>%
  filter(n() > 5) %>%
  droplevels()

new_car_outliers_removed <- new_car_outliers_removed  %>%
  group_by(year) %>%
  filter(n() > 5) %>%
  droplevels()

new_car_outliers_removed <- new_car_outliers_removed  %>%
  group_by(state) %>%
  filter(n() > 5) %>%
  droplevels()

```


#Regression analysis
```{r}
lm_before_remove_outliers <- lm(price ~., data = new_car_collapsed_knn)
summary(lm_before_remove_outliers)
lm_before_remove_outliers %>% glance()
tidy(lm_before_remove_outliers)

lm_after_remove_outliers <- lm(price ~., data = new_car_outliers_removed)
summary(lm_after_remove_outliers)
lm_after_remove_outliers %>% glance()
tidy(lm_after_remove_outliers)
```

#Summary of the comparison of the regression model with and without outliers 
```{r}
compare_before <- list(glance(lm_before_remove_outliers))
compare_after <- list(glance(lm_after_remove_outliers))

compare <- bind_rows(compare_before, compare_after)
compare1 <- as.data.frame(compare)
rownames(compare1) <- c("Before", "After")
compare1
```

#Plotting the residuals
```{r}
ggplot(lm_after_remove_outliers, aes(x=fitted(lm_after_remove_outliers), y=residuals(lm_after_remove_outliers))) +
  geom_point() +
  labs(x= "fitted", y = "residuals") +
  geom_hline(yintercept = 0, linetype = "dashed")
```

#Predicting Linear Models
```{r}
set.seed(12)
trainingRowIndex <- sample(1:nrow(new_car_outliers_removed), 0.8*nrow(new_car_outliers_removed))
trainingData <- new_car_outliers_removed[trainingRowIndex, ]
testData <- new_car_outliers_removed[-trainingRowIndex, ]

lmMod <- lm(price ~., data= trainingData)
PricePred <- predict(lmMod, newdata= testData, interval = "confidence")

summary(lmMod)
```

#Accuracy of the Predicting model
```{r}
actuals_preds <- data.frame(cbind(actuals=testData$price, PricePred)) 

head(actuals_preds)
```

#Conclusion
In order to perform the regression analysis on the dataset, we have to first remove the very low count levels of factors.
The regression model with outliers removed have a standard error deviation of 5499. 
The coefficient of determination is 0.7259 and the adjusted r-square is 0.7056.
The F-statistic is 35.71
